Manuscript_Github_Pchabaudi_pir_210105

### 2021-03-12 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd

```{r SorL_stacked_eachGene_fig1b}
  tpm_data <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = Geneid) %>% 
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

@@

(
  y <- ggplot(data = tpm_data_line, 
              aes(x = stage, 
                  y = total_tpm_median, 
                  fill = experiment, 
                  group = experiment)) +
    geom_bar(position = position_dodge2(preserve = 'single'),
             stat = 'identity') +
    scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
    geom_errorbar(aes(x = stage,
                      ymax = total_tpm_max, 
                      ymin = total_tpm_min),
                  position = position_dodge2(preserve = 'single'),
                  size = 0.25) +
    scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_text(angle = 90, 
                                   size = 6, 
                                   vjust = 0.5),
          axis.text.y = element_text(size = 5),
          axis.title.y = element_text(size = 8),
          legend.title = element_text(size = 7),
          legend.text = element_text(size = 5),
          legend.key.size = unit(3,'mm'),
          legend.key.width = unit(2,'mm')) +
    scale_size(name = 'Number of genes detected over threshold', 
               range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) +
    facet_wrap_custom(~Geneid, 
                      scales = 'free_y',
                      scale_overrides = list(
                        scale_override(3, scale_y_continuous(breaks = seq(0,650,length.out = 11))),
                        scale_override(2, scale_y_continuous(breaks = seq(0,900,length.out = 11))),
                        scale_override(1, scale_y_continuous(breaks = seq(0,310,length.out = 11)))
                      )) +
    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)
)
#Could also use functions from a package called 'lemon' to add the lines to the top two facets.
```
```

>>>>

```{r SorL_stacked_eachGene_fig1b}
  tpm_data <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = Geneid) %>% 
    mutate(tpm = ifelse(tpm > 1, tpm, 0)) %>% #Filtering step for low value pir genes.
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

@@

(
  y <- ggplot(data = tpm_data_line, 
              aes(x = stage, 
                  y = total_tpm_median, 
                  fill = experiment, 
                  group = experiment)) +
    geom_bar(position = position_dodge2(preserve = 'single'),
             stat = 'identity') +
    scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
    geom_errorbar(aes(x = stage,
                      ymax = total_tpm_max, 
                      ymin = total_tpm_min),
                  position = position_dodge2(preserve = 'single'),
                  size = 0.25) +
    scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_text(angle = 90, 
                                   size = 6, 
                                   vjust = 0.5),
          axis.text.y = element_text(size = 5),
          axis.title.y = element_text(size = 8),
          legend.title = element_text(size = 7),
          legend.text = element_text(size = 5),
          legend.key.size = unit(3,'mm'),
          legend.key.width = unit(2,'mm')) +
    scale_size(name = 'Number of genes detected over threshold', 
               range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) +
    facet_wrap_custom(~Geneid, 
                      scales = 'free_y',
                      scale_overrides = list(
                        scale_override(3, scale_y_continuous(breaks = seq(0,650,length.out = 11))),
                        scale_override(2, scale_y_continuous(breaks = seq(0,900,length.out = 11))),
                        scale_override(1, scale_y_continuous(breaks = seq(0,310,length.out = 11)))
                      )) +
    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)
)
#Could also use functions from a package called 'lemon' to add the lines to the top two facets.
```

### 2021-03-12 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd

@@

                       if(
                         identical(list_element,
                                   first(zscore_list)) & is.character(annotation_clade)
                       ){
                         clade_ha <- rowAnnotation(
                           'subfamily' = anno_simple(
                             clade_info,
                             col = annotation_col
                           )
                         )
                       }

@@

                       h = Heatmap(heatmap_data,
                                   name = paste0('z-score',
                                                 extra_title),
                                   col = c(cbPalette[6],
                                           'white', 
                                           cbPalette[7]),
                                   row_names_gp = gpar(fontsize = ifelse(
                                     nrow(heatmap_data) > 13, 
                                     12*(13/nrow(heatmap_data)), 
                                     12)
                                   ),
                                   column_names_gp = gpar(fontsize = 10),
                                   row_dend_width = unit(3, "cm"),
                                   right_annotation = if(tpm_mat){
                                     tpm_ha
                                   }else{
                                     NULL
                                   },
                                   left_annotation = 
                                     if(identical(list_element,
                                                  first(zscore_list)) & is.character(annotation_clade)){
                                       clade_ha
                                     }else{
                                       NULL
                                     }, 
                                   #Only want the clade annotation when it is first heatmap
                                   width = unit(4.5, 'cm'),
                                   cluster_columns = FALSE,
                                   height = unit(7,'cm'),
                                   row_title = NULL,
                                   cluster_rows = FALSE,
                                   column_title = paste(title),
                                   heatmap_legend_param = list(
                                     at = c(3,2,1,0,-1,-2,-3),
                                     labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                   ),
                                   ...
                       )
                       return(h)
                     })

@@

```
## Figure 3b - Highest expressed all _pir_ heatmap

@@

@@

         # combined_data_ordered <- melt(
         #   (
         #     combined_data_pir %>% 
         #       rownames_to_column(var = 'Geneid')
         #   ),
         #   id = 'Geneid',
         #   variable.name = 'stage',
         #   value.name = 'tpm') %>% 
         #   group_by(Geneid) %>% 
         #   top_n(tpm, 
         #         n = 1) %>% 
         #   arrange(stage,
         #           desc(tpm)) %>%
         #   distinct(Geneid, 
         #            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
         #   filter(tpm > 0)

@@

                             Heatmap(
                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
                               cluster_rows = FALSE,
                               cluster_columns = FALSE,
                               show_row_names = FALSE,
                               name = 'log(TPM+1)', 
                               col = colorRamp2(c(0,
                                                  maxcol/4,
                                                  maxcol/2,
                                                  3*maxcol/4,
                                                  maxcol),
                                                viridis(5)),
                               na_col = 'white', 
                               border = TRUE,
                               border_gp = gpar(lwd = 0.1),
                               column_names_gp = gpar(fontsize = 10),
                               row_dend_width = unit(3, "cm"),
                               width = unit(4.5, 'cm'),
                               height = unit(7,'cm'),
                               row_title = NULL,
                               left_annotation = if(
                                 list_name %in% first(names(combined_data_pir_ht_list))
                               ){
                                 clade_ha
                               }else{
                                 NULL
                               },
                               right_annotation = if(
                                 list_name %in% last(names(combined_data_pir_ht_list))
                               ){
                                 anc_ha
                               }else{
                                 NULL
                               },
                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
                               #                                        cir_info$Geneid)],
                               column_title = paste(list_name),
                               heatmap_legend_param = list(
                                 labels_gp = gpar(fontsize = 10),
                                 title_gp = gpar(fontsize = 10,
                                                 fontface = 'bold'),
                                 at = c(0,
                                        maxcol/4,
                                        maxcol/2,
                                        3*maxcol/4,
                                        maxcol)
                               )
                             )
                           }
         )
         ht_list <- Reduce(`+`, ht_list)
         clade_lgd <- Legend(
           title = 'sub-family',
           at = names(SorL_col),
           legend_gp = gpar(fill = SorL_col)
         )

>>>>

```{r function_ordering}
ordering <- function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}

order_vector <- c("Early.Rings",
                  "Early.Mid.Rings",
                  "Late.Mid.Rings",
                  "Late.Rings",
                  "Rings.Troph.conversion",
                  "Early.Trophs",
                  "Mid.Trophs",
                  "Late.Trophs")
```

@@

                       if(
                         identical(list_element,
                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
                       ){
                         clade_ha <- rowAnnotation(
                           'subfamily' = anno_simple(
                             clade_info,
                             col = annotation_col
                           )
                         )
                       }

@@

                       h = Heatmap(heatmap_data,
                                   name = paste0('z-score',
                                                 extra_title),
                                   col = c(cbPalette[6],
                                           'white', 
                                           cbPalette[7]),
                                   row_names_gp = gpar(fontsize = ifelse(
                                     nrow(heatmap_data) > 13, 
                                     12*(13/nrow(heatmap_data)), 
                                     12)
                                   ),
                                   column_names_gp = gpar(fontsize = 10),
                                   row_dend_width = unit(3, "cm"),
                                   right_annotation = if(tpm_mat){
                                     tpm_ha
                                   }else{
                                     NULL
                                   },
                                   left_annotation = 
                                     if(identical(list_element,
                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
                                       clade_ha
                                     }else{
                                       NULL
                                     }, 
                                   #Only want the clade annotation when it is first heatmap
                                   width = unit(4.5, 'cm'),
                                   cluster_columns = FALSE,
                                   height = unit(7,'cm'),
                                   row_title = NULL,
                                   cluster_rows = FALSE,
                                   column_title = paste(title),
                                   heatmap_legend_param = list(
                                     at = c(3,2,1,0,-1,-2,-3),
                                     labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                   ),
                                   ...
                       )
                       return(h)
                     })

@@

## Figure 3b - Highest expressed all _pir_ heatmap

@@

num_SorL <- cir_info %>% 
  group_by(SorL) %>% 
  summarise(num = n_distinct(Gene.ID))

labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
                        num_SorL$SorL)

@@

@@

                             Heatmap(
                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
                               cluster_rows = FALSE,
                               cluster_columns = FALSE,
                               show_row_names = FALSE,
                               name = 'log(TPM+1)', 
                               col = colorRamp2(c(0,
                                                  maxcol/4,
                                                  maxcol/2,
                                                  3*maxcol/4,
                                                  maxcol),
                                                viridis(5)),
                               na_col = 'white', 
                               border = TRUE,
                               border_gp = gpar(lwd = 0.1),
                               column_names_gp = gpar(fontsize = 10),
                               row_dend_width = unit(3, "cm"),
                               width = unit(4.5, 'cm'),
                               height = unit(7,'cm'),
                               row_title = NULL,
                               left_annotation = if(
                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
                               ){
                                 clade_ha
                               }else{
                                 NULL
                               },
                               right_annotation = if(
                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
                               ){
                                 anc_ha
                               }else{
                                 NULL
                               },
                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
                               #                                        cir_info$Geneid)],
                               column_title = paste(list_name),
                               heatmap_legend_param = list(
                                 labels_gp = gpar(fontsize = 10),
                                 title_gp = gpar(fontsize = 10,
                                                 fontface = 'bold'),
                                 at = c(0,
                                        maxcol/4,
                                        maxcol/2,
                                        3*maxcol/4,
                                        maxcol)
                               )
                             )
                           }
         )
         ht_list <- Reduce(`+`, ht_list)
         clade_lgd <- Legend(
           title = 'sub-family',
           at = labels_SorL,
           legend_gp = gpar(fill = SorL_col)
         )
